---
title: "Mini Data-Analysis Deliverable 1"
output: github_document
---

# Welcome to your (maybe) first-ever data analysis project!

And hopefully the first of many. Let's get started:

1.  Install the [`datateachr`](https://github.com/UBC-MDS/datateachr) package by typing the following into your **R terminal**:

<!-- -->

    install.packages("devtools")
    devtools::install_github("UBC-MDS/datateachr")

2.  Load the packages below.

```{r}
library(datateachr)
library(tidyverse)
```

3.  Make a repository in the <https://github.com/stat545ubc-2023> Organization. You can do this by following the steps found on canvas in the entry called [MDA: Create a repository](https://canvas.ubc.ca/courses/126199/pages/mda-create-a-repository). One completed, your repository should automatically be listed as part of the stat545ubc-2023 Organization.

# Instructions

## For Both Milestones

-   Each milestone has explicit tasks. Tasks that are more challenging will often be allocated more points.

-  Each milestone will be also graded for reproducibility, cleanliness, and coherence of the overall Github submission. 

- While the two milestones will be submitted as independent deliverables, the analysis itself is a continuum - think of it as two chapters to a story. Each chapter, or in this case, portion of your analysis, should be easily followed through by someone unfamiliar with the content. [Here](https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R/) is a good resource for what constitutes "good code". Learning good coding practices early in your career will save you hassle later on!

- The milestones will be equally weighted. 

## For Milestone 1

**To complete this milestone**, edit [this very `.Rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-1.Rmd) directly. Fill in the sections that are tagged with `<!--- start your work below --->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an `.md` file by changing the YAML output settings from `output: html_document` to `output: github_document`. Commit and push all of your work to the mini-analysis GitHub repository you made earlier, and tag a release on GitHub. Then, submit a link to your tagged release on canvas.

**Points**: This milestone is worth 36 points: 30 for your analysis, and 6 for overall reproducibility, cleanliness, and coherence of the Github submission. 

# Learning Objectives

By the end of this milestone, you should:

-   Become familiar with your dataset of choosing
-   Select 4 questions that you would like to answer with your data
-   Generate a reproducible and clear report using R Markdown
-   Become familiar with manipulating and summarizing your data in tibbles using `dplyr`, with a research question in mind.

# Task 1: Choose your favorite dataset 

The `datateachr` package by Hayley Boyce and Jordan Bourak currently composed of 7 semi-tidy datasets for educational purposes. Here is a brief description of each dataset:

-   *apt_buildings*: Acquired courtesy of The City of Toronto's Open Data Portal. It currently has 3455 rows and 37 columns.

-   *building_permits*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 20680 rows and 14 columns.

-   *cancer_sample*: Acquired courtesy of UCI Machine Learning Repository. It currently has 569 rows and 32 columns.

-   *flow_sample*: Acquired courtesy of The Government of Canada's Historical Hydrometric Database. It currently has 218 rows and 7 columns.

-   *parking_meters*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 10032 rows and 22 columns.

-   *steam_games*: Acquired courtesy of Kaggle. It currently has 40833 rows and 21 columns.

-   *vancouver_trees*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 146611 rows and 20 columns.

**Things to keep in mind**

-   We hope that this project will serve as practice for carrying our your own *independent* data analysis. Remember to comment your code, be explicit about what you are doing, and write notes in this markdown document when you feel that context is required. As you advance in the project, prompts and hints to do this will be diminished - it'll be up to you!

-   Before choosing a dataset, you should always keep in mind **your goal**, or in other ways, *what you wish to achieve with this data*. This mini data-analysis project focuses on *data wrangling*, *tidying*, and *visualization*. In short, it's a way for you to get your feet wet with exploring data on your own.

And that is exactly the first thing that you will do!

1.1 **(1 point)** Out of the 7 datasets available in the `datateachr` package, choose **4** that appeal to you based on their description. Write your choices below:

**Note**: We encourage you to use the ones in the `datateachr` package, but if you have a dataset that you'd really like to use, you can include it here. But, please check with a member of the teaching team to see whether the dataset is of appropriate complexity. Also, include a **brief** description of the dataset here to help the teaching team understand your data.

<!-------------------------- Start your work below ---------------------------->
> Here, I have only chosen 4 from the given 7, hence no explaination about the datasets is done. 
>1: *building_permits*\
>2: *cancer_sample*\
>3: *steam_games*\
>4: *vancouver_trees*\

<!----------------------------------------------------------------------------->

1.2 **(6 points)** One way to narrowing down your selection is to *explore* the datasets. Use your knowledge of dplyr to find out at least *3* attributes about each of these datasets (an attribute is something such as number of rows, variables, class type...). The goal here is to have an idea of *what the data looks like*.

*Hint:* This is one of those times when you should think about the cleanliness of your analysis. I added a single code chunk for you below, but do you want to use more than one? Would you like to write more comments outside of the code chunk?

<!-------------------------- Start your work below ---------------------------->

```{r}
# I wanted to check the classes of the datasets in order to understand which functions I need to use during my data manipulation steps.
# spec_tbl_df is basically a wrapped tbl_df, one can subset the sec_tbl_df and convert it to a tbl_df.
class(building_permits)
class(cancer_sample)
class(steam_games)
class(vancouver_trees) 

# Glimpse gives rows x columns, variables and some of the data as well with the columns' data type, pretty useful
# I use these mainly to see the variables, NA values, and data types of the variables. I wanted to extract the data from columns so I needed to see how and what they are stored as.
glimpse(building_permits)
glimpse(cancer_sample)
glimpse(steam_games)
glimpse(vancouver_trees)

# For the building_permits dataset, I wanted to check the counts of unique ids and unique permit numbers. 
# To obtain a healthier comparison, I mutated the permit_number column in a way that the dashes in the permit numbers, which can contain same number (some of them contain dates) but with dashes and can create ambiguity, are eliminated. 
building_permits_new <- building_permits %>% mutate(permit_number_dashless = gsub("-", "", permit_number))
building_permits_new %>% summarise(num_of_rows=n(), num_of_ids=length(unique(bi_id)), num_of_permit_nums=length(unique(permit_number_dashless))) # everything seems equal :D

# For the cancer_sample dataset, I wanted to check the unique id count only since the other things can be coincidentally the same
cancer_sample %>% summarise(num_of_rows=n(), num_of_ids=length(unique(ID))) # everything seems fine

# For the steam_games dataset, I wanted to check the counts of unique ids and unique urls, since urls must be unique as well due to internet protocols
steam_games %>% summarize(num_of_rows=n(), num_of_ids=length(unique(id)), num_of_urls=length(unique(url))) # it is all the same, which is a positive sign

# For the vancouver_trees dataset, I wanted to check the counts of unique ids and unique locations (basically unique longitude and latitude pairs) to see if there are multiple entries under the same id and same location
unq_vanc_trees_locs <- unique(select(vancouver_trees, c(longitude,latitude))) # select only the unique location pairs
 vancouver_trees %>% summarize(num_of_rows=n(), num_of_ids=length(unique(tree_id)), num_of_locations=dim(unq_vanc_trees_locs)[1]) # data shows every entry has its unique id but not unique location which may mean some of the trees are either parasites or colliding with another one. The second option is bizarre.

```

<!----------------------------------------------------------------------------->

1.3 **(1 point)** Now that you've explored the 4 datasets that you were initially most interested in, let's narrow it down to 1. What lead you to choose this one? Briefly explain your choice below.

<!-------------------------- Start your work below ---------------------------->
>Despite *vancouver_trees* dataset, the datasets were consistent in terms of their entry ids and other unique variables. By inspection, it can be seen that *building_permits* dataset contains too many NA values. On the other hand, *cancer_sample* dataset contains so many variables and these variables can be coexpressing different things on the diagnosis as well, which is a hard-to-detect relation. Hence, at the end, I decided on moving with *steam_games* dataset, which seemed like the most balanced one among them. My pandemic was all about video games, so it is interesting to see the data of the steam games to me as well!
<!----------------------------------------------------------------------------->

1.4 **(2 points)** Time for a  final decision! Going back to the beginning, it's important to have an *end goal* in mind. For example, if I had chosen the `titanic` dataset for my project, I might've wanted to explore the relationship between survival and other variables. Try to think of 1 research question that you would want to answer with your dataset. Note it down below.

<!-------------------------- Start your work below ---------------------------->
> What is the relationship between the review count and the review score for the games? Review score is the one that is given in a percentage (94% liked the game for example).
<!----------------------------------------------------------------------------->

# Important note

Read Tasks 2 and 3 *fully* before starting to complete either of them. Probably also a good point to grab a coffee to get ready for the fun part!

This project is semi-guided, but meant to be *independent*. For this reason, you will complete tasks 2 and 3 below (under the **START HERE** mark) as if you were writing your own exploratory data analysis report, and this guidance never existed! Feel free to add a brief introduction section to your project, format the document with markdown syntax as you deem appropriate, and structure the analysis as you deem appropriate. If you feel lost, you can find a sample data analysis [here](https://www.kaggle.com/headsortails/tidy-titarnic) to have a better idea. However, bear in mind that it is **just an example** and you will not be required to have that level of complexity in your project.

# Task 2: Exploring your dataset 

If we rewind and go back to the learning objectives, you'll see that by the end of this deliverable, you should have formulated *4* research questions about your data that you may want to answer during your project. However, it may be handy to do some more exploration on your dataset of choice before creating these questions - by looking at the data, you may get more ideas. **Before you start this task, read all instructions carefully until you reach START HERE under Task 3**.

2.1 **(12 points)** Complete *4 out of the following 8 exercises* to dive deeper into your data. All datasets are different and therefore, not all of these tasks may make sense for your data - which is why you should only answer *4*. 

Make sure that you're using dplyr and ggplot2 rather than base R for this task. Outside of this project, you may find that you prefer using base R functions for certain tasks, and that's just fine! But part of this project is for you to practice the tools we learned in class, which is dplyr and ggplot2.

1.  Plot the distribution of a numeric variable.
2.  Create a new variable based on other variables in your data (only if it makes sense)
3.  Investigate how many missing values there are per variable. Can you find a way to plot this?
4.  Explore the relationship between 2 variables in a plot.
5.  Filter observations in your data according to your own criteria. Think of what you'd like to explore - again, if this was the `titanic` dataset, I may want to narrow my search down to passengers born in a particular year...
6.  Use a boxplot to look at the frequency of different observations within a single variable. You can do this for more than one variable if you wish!
7.  Make a new tibble with a subset of your data, with variables and observations that you are interested in exploring.
8.  Use a density plot to explore any of your variables (that are suitable for this type of plot).

2.2 **(4 points)** For each of the 4 exercises that you complete, provide a *brief explanation* of why you chose that exercise in relation to your data (in other words, why does it make sense to do that?), and sufficient comments for a reader to understand your reasoning and code.

<!-------------------------- Start your work below ---------------------------->
## Introduction
[Steam](https://store.steampowered.com) is a platform that maintains and sells video games for computers. _steam\_games_ dataset contains a subset of games from Steam platform. It contains relevant information about each game such as its original or discounted price, name, url, reviews, etc. I will be analysing this dataset by my selection of questions from the 8 questions mentioned above. Furthermore, the data analysis is done by _ggplot2_ and _deplyr_ libraries. 

1.  Plot the distribution of a numeric variable.

> To invest in something, one first needs to do a market research and be informed about the general prices. Thus, it makes sense to see the general distribution of the original prices of the games in Steam and I will color the different types of products accordingly. I first tried to not scale the axes but since the count of some pricepoints and the price diversity are wide and do not produce a reasonable and understandable histogram, I wanted to condense them by using log10 scaling. However, some games are free which means their transformation would converge. To prevent the convergence of these data, I have added a small price (0.001) to their original price. I did not do anything abour 0 count for a price point since it should not be represented on the graph anyway. The warnings indicate 0 count bins which is fine and nothing concerning. I also tried to plot the density function, but it does not represents well, so I am keeping the histogram.

```{r}
# since I will use log scale, I first add a small amount (0.001) to the every original_price value before using log function, just to prevent 0 values from converging
anti_converge_steam_games = steam_games %>% mutate(original_price = original_price + 0.001) %>% filter(!is.na(original_price))

# Here, I also categorized the games due to their types (if the game is a bundle, app, sub, or NA) to see the difference between their price average as well compared to other types
# I scale axes to make the graph more legible and interpretable since the price interval is too wide and app count dominates the type comparison as well.
ggplot(anti_converge_steam_games, aes(original_price)) +
  geom_histogram( aes(fill=types), bins=60)+ # I have used bins as 60, since it was the best option for the sake of reading and interpretation
  scale_x_log10("Original Price (Scale: log10)") +
  scale_y_log10("Count (Scale: log10)") +
  ggtitle("Histogram for the count distribution of games on original price") 

```

3.  Investigate how many missing values there are per variable. Can you find a way to plot this?

> When visually inspecting the data with glimpse() function, I saw some NA values for the achievement count, original price and discount price. It is important to know how many NA values are there for these variables in order to conduct a proper data analysis. Furthermore, if they contain too many missing information, it is not so healthy to conduct analysis on them for the next questions and MDS2. It seems like some data has too much NA values. For example, I would like to engage discount_price as well but it seems like it would not be a representative analysis, at least on that dataset.

```{r}
# Here, I  basically count the number of na values across every column and put them into a summarise table - is.na returns 1 if it is NA and we add 1 onto the sum which allows us to count NA occurrances
steam_games_na <- steam_games %>% summarise(across(everything(),~sum(is.na(.))))
col_names <- colnames(steam_games_na) # storing variable names into col_names 
tmp_df <- t(steam_games_na) # transposing the columns into rows or variables into values
colnames(tmp_df) <- c("values") # giving the column name as "values" that contains the counts
steam_games_na <- as_tibble(tmp_df) # converting summarise table to a tibble
steam_games_na["col_names"] <- col_names # appending column names to corresponding counts as a new variable named "col_names"

# I feel like it is better to have a 90 degree flipped histogram is a better option to represent the counts of NA for each column, since the names are long and a lot, it felt more intuitive and easy to read
ggplot(steam_games_na, aes(col_names, values)) + 
  geom_col(fill="blue") +
  xlab("Column Names") +
  ylab("Count of NA entries") +
  ggtitle("Histogram for the count of NA entries for each variable") +
  coord_flip()
```

4.  Explore the relationship between 2 variables in a plot.

> As mentioned on my research question, I wonder if there is any relation between the review score and the review count. So, I will investigate the relation between the score and the number of the reviews given. In order to do that, I will extract the score and review count information from all_reviews and store them in a new columns. I believe, lower count in the reviews are a bit open to incline on either to have a really good score or a really bad score. This time, review counts were too crowded and dense, in order to create a better representation which allows us to see the different scores and review type and count, I scaled x axis with log10. On steam games dataset, if a game has received reviews smaller than 10 times, its review type cannot be decided. So, I filtered those with <10 reviews. Now, log10 cannot converge since the minimum count is 10/the data does not contain 0 review count games. This plot shows review scores, review counts, and review types since I have colored the dots according to the review type. It seems like people tend to give more reviews that have higher enjoyability. 
 
```{r}
augmented_steam_games <- steam_games %>% mutate(score=as.integer(gsub("%", "", str_extract(pattern="\\d+%", string=all_reviews))), # extracting the score out of 100 of the reviews for a game
                                                review_type=gsub(",", "", str_extract(pattern="(\\w+,)|(\\w+ \\w+,)", string=all_reviews)), # extracting the review type/average of a game (negative, positive etc.)
                                                review_cnt=as.integer(gsub(",", "",str_extract(pattern="(\\d+,\\d+)|(\\d+)", string=all_reviews ))))# extracting the all review count of a game

augmented_steam_games <- augmented_steam_games %>% filter(review_type!="user reviews") # filtering these ones since they need more reviews to be categorized into some category (they have review count smaller than 10)

# Here, I plot a jitter graph that shows the score vs review count and colors the dots according to their review type.
ggplot(augmented_steam_games) +
  geom_jitter(aes(x=review_cnt, y=score, colour=review_type), size=0.05 ) +
  scale_x_log10("Review Count (Scaled: log10)", labels=scales::number_format() )+ # that does not trigger any divergence since a game has to have at least 10 reviews to be categorized under a review type
  ggtitle("Jitter plot that shows score vs review count") +
  ylab("Score")
```

5.  Filter observations in your data according to your own criteria. Think of what you'd like to explore - again, if this was the `titanic` dataset, I may want to narrow my search down to passengers born in a particular year...

> I was curious about if the satisfaction of the customers are somewhat related to the price since people tend to expect less from low priced games. I will only check the prices of the "Very Positive" graded games (under all reviews column). So, one can notice if people say "it is a good game" only if that game is not expensive. I scaled y axis/count of the pricepoint since some counts dominated which made small counts invisible and I wanted to see the small counts in the graph as well in order to have a more comprehensive analysis. In order to prevent converging in 0 dollar games. So, I have used *anti_convergence_games* tibble that I have created in question 1 of this section. I provide a boxplot which shows the outlier and average a bit more conveniently and a histogram that shows the bin distribution better. It seems like people tend to like cheaper games. And people only voted apps as "very positive" according to the boxplot. Again, I have scaled the y axis with log10 in order to see the details better. The warnings are due to 0 counts which would not be represented anyway. Density plot was not detailed enough again, so I am keeping the histogram instead.

```{r}
# Here, I only filter in the one that contain "Very Positive" in their all_reviews field.
very_positive_games <- anti_converge_steam_games %>% filter(grepl(pattern="Very Positive", x=all_reviews, ignore.case=TRUE))

# That is a boxplot which shows the original price distribution summary of very positive games, furthermore x axis is according to the product type (app, bundle etc.)
ggplot(very_positive_games) +
  geom_boxplot(aes(x=types,y=original_price), color="purple") +
  scale_y_log10("Original Price (Scale: log10)") +
  ggtitle("Boxplot that shows ") +
  xlab("Product Type") 
  
# This is the histogram that shows the counts of original price points of the very positive games
ggplot(very_positive_games) + 
  geom_histogram(aes(original_price), color="purple", bins=40) +
  scale_y_log10("Count (Scale: log10)", labels = scales::number_format()) +
  xlab("Original Price") +
  scale_x_continuous(labels=scales::dollar_format())
```

<!----------------------------------------------------------------------------->

# Task 3: Choose research questions 

**(4 points)** So far, you have chosen a dataset and gotten familiar with it through exploring the data. You have also brainstormed one research question that interested you (Task 1.4).  Now it's time to pick 4 research questions that you would like to explore in Milestone 2! Write the 4 questions and any additional comments below. 

<!--- *****START HERE***** --->
1. Is there a correlation between the game price and count of the reviews of that game? If yes, how are they correlated?
2. Is there a correlation between the game price and the overall review score of that game? If yes, how are they correlated?
3. What is the distribution of genres of the games? Is the distribution correlated with the price in a way? 
4. What is the distribution of the languages of the games? Is the distribution correlated with the review score of that game in any sense?

<!----------------------------->

# Overall reproducibility/Cleanliness/Coherence Checklist 

## Coherence (0.5 points)

The document should read sensibly from top to bottom, with no major continuity errors. An example of a major continuity error is having a data set listed for Task 3 that is not part of one of the data sets listed in Task 1. 

## Error-free code (3 points) 

For full marks, all code in the document should run without error. 1 point deduction if most code runs without error, and 2 points deduction if more than 50% of the code throws an error. 

## Main README (1 point)

There should be a file named `README.md` at the top level of your repository. Its contents should automatically appear when you visit the repository on GitHub.

Minimum contents of the README file:

-   In a sentence or two, explains what this repository is, so that future-you or someone else stumbling on your repository can be oriented to the repository.
-   In a sentence or two (or more??), briefly explains how to engage with the repository. You can assume the person reading knows the material from STAT 545A. Basically, if a visitor to your repository wants to explore your project, what should they know?

Once you get in the habit of making README files, and seeing more README files in other projects, you'll wonder how you ever got by without them! They are tremendously helpful.

## Output (1 point)

All output is readable, recent and relevant:

-   All Rmd files have been `knit`ted to their output md files. 
-   All knitted md files are viewable without errors on Github. Examples of errors: Missing plots, "Sorry about that, but we can't show files that are this big right now" messages, error messages from broken R code
-   All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
-   There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted.

(0.5 point deduction if any of the above criteria are not met. 1 point deduction if most or all of the above criteria are not met.)

Our recommendation: right before submission, delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant. Then, after your final commit and push to Github, CHECK on Github to make sure that everything looks the way you intended!

## Tagged release (0.5 points)

You've tagged a release for Milestone 1. 

### Attribution

Thanks to Icíar Fernández Boyano for mostly putting this together, and Vincenzo Coia for launching.
